---
title: "Looking for better tips? Go to... "
author: "Andrea Celleri "
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

knitr::opts_chunk$set(fig.width = 7)
knitr::opts_chunk$set(fig.height = 4)


library(readr)
library(ggplot2)
library(dplyr)
library(ggmap)
library(viridis)
library(forcats)
library(smodels)
```


```{r}
taxi <- read_csv("https://statsmaths.github.io/stat_data/nyc_taxi_50.csv")
nta_code <- read_csv("https://raw.githubusercontent.com/toddwschneider/nyc-taxi-data/master/data/taxi-zone-lookup-with-ntacode.csv")

```


```{r}

taxi$tip_percent <- taxi$tip_amount / taxi$fare_amount
taxi <- filter(taxi, tip_percent < 2)

nyc_cord <- filter(taxi, dropoff_longitude > -75, dropoff_longitude < -65, dropoff_latitude < 42, dropoff_latitude > 40  )

taxi_nta_pickup <-left_join(taxi, nta_code, by = c("pickup_NTACode" = "ntacode"))

taxi_nta <- left_join(taxi_nta_pickup, nta_code, by= c("dropoff_NTACode" = "ntacode"), suffix = c("_pickup", "_dropoff"), tip_percent < 2)

arrival_laguardia <- filter(taxi_nta, zone_pickup == "LaGuardia Airport" ,pickup_longitude > -74.8, pickup_longitude < -73.2, pickup_latitude < 42, pickup_latitude > 40.76)

departure_laguardia_mnh <- filter(taxi_nta, dropoff_NTACode == "QN98",zone_dropoff == "LaGuardia Airport" , dropoff_longitude > -74.8, dropoff_longitude < -73.2, dropoff_latitude < 42, dropoff_latitude > 40.76 , borough_pickup =="Manhattan")

departure_laguardia <- filter(taxi_nta, dropoff_NTACode == "QN98",zone_dropoff == "LaGuardia Airport" , dropoff_longitude > -74.8, dropoff_longitude < -73.2, dropoff_latitude < 42, dropoff_latitude > 40.76 )

```

```{r}
arrival_laguardia_mnh <- filter(arrival_laguardia, borough_dropoff =="Manhattan")

zone_pickup_summary <- group_summarize(arrival_laguardia_mnh, zone_dropoff)

arrivals_laguardia_uptown <- filter(arrival_laguardia_mnh, zone_dropoff %in% c( "Marble Hill", "Inwood", "Washigton Heights", "West Harlem", "Hamilton Heights", "Manhattanville", "Morningside Heights", "Central Harlem", "Harlem", "Astor Row", "Sugar Hill", "Marcus Garvey Park", "Le Petit Senegal", "East Harlem", "Upper East Side", "Lenox Hill", "Carnegie Hill", "Yorkville", "Upper West Side", "Lincoln Square", "Morningside Heights", "Washington Heights North", "Washington Heights South", "Inwood", "Inwood Hill Park", "Highbridge Park", "Hamilton Heights", "Central Harlem North", "East Harlem North", "Morningside Heights"))

zone_pickup_summary_uptown <- group_summarize(arrivals_laguardia_uptown, zone_dropoff)


arrivals_laguardia_midtown <- filter(arrival_laguardia_mnh, zone_dropoff %in% c("Midtown", "Columbus Circle", "Sutton Place", "Rickefeller Center", "Diamond District", "Trutle Bay", "Midtown East", "Midtown", "Tudon City", "Little Brazil" , "Hudson Yards", "Midtown West", "Garment District", "Koreatown", "Murray Hill", "Tenderloin", "Madison Sq", "Clinton West", "East Chelsea", "Flatiron", "Garment District", "Union sq" , "Kips Bay", "Murray Hill", "West Chelsea/Hudson Yards", "Sutton Place/Turtle Bay North", "UN/Turtle Bay South","Sutton Place/Turtle Bay North", "Penn Station/Madison Sq West", "Times Sq/Theatre District", "Gramercy", "Union Sq", "Midtown North", "Midtown South"))


zone_pickup_summary_midtown <- group_summarize(arrivals_laguardia_midtown, zone_dropoff)

arrivals_laguardia_downtown <- filter(arrival_laguardia_mnh, zone_dropoff %in% c("East Village", "Alphabet City", "TriBeCa/Civic Center", " Little Italy/NoLiTa", "SoHo", "Hudson Sq", " Lower East Side", "Chinatown", "seaport","Two Bridges/Seward Park", "World Trade Center", "Financial District South", "Financial District North", "Battery Park City", "Battery Park", "Greenwich Village North", "Greenwich Village South", "Meatpacking/West Village West", "West Village", "Seaport", "Lower East Side", "Little Italy/NoLiTa"))

zone_pickup_summary_downtown <- group_summarize(arrivals_laguardia_downtown, zone_dropoff)

arrivals_laguardia_uppereast <- filter(arrival_laguardia_mnh, zone_dropoff %in% c("Yorkville East", "Upper East Side North", "Upper East Side South", "Yorkville West", "East Harlem South", "Lenox Hill West", "Lenox Hill East"))

zone_pickup_summary_uppereast <- group_summarize(arrivals_laguardia_uppereast, zone_dropoff)

arrivals_laguardia_upperwest <- filter(arrival_laguardia_mnh, zone_dropoff %in% c( "Lincoln Square East", "Lincoln Square West","Bloomingdale",  "Upper West Side North", "Upper West Side South", "Manhattan Valley", "Bloomingdale"))

zone_pickup_summary_upperewest <- group_summarize(arrivals_laguardia_upperwest, zone_dropoff)

```

```{r}
taxi_nta_mnh <- filter(taxi_nta, borough_dropoff == "Manhattan", tip_percent < Inf, tip_percent < 2)

taxi_nta_queens <- filter(taxi_nta, borough_dropoff == "Queens", tip_percent < Inf, tip_percent < 2)
taxi_nta_queens_summary <- group_summarize(taxi_nta_queens, zone_dropoff)

taxi_nta_brk <- filter(taxi_nta, borough_dropoff == "Brooklyn", tip_percent < Inf, tip_percent < 2)
taxi_nta_brk_summary <- group_summarize(taxi_nta_brk, zone_dropoff)



taxi_nta <- filter(taxi_nta, tip_percent < Inf)

taxi_nta_mnh_summary <- group_summarize(taxi_nta_mnh, zone_dropoff)



taxi_nta_mnh_uptown <- filter(taxi_nta_mnh, zone_dropoff %in% c( "Marble Hill", "Inwood", "Washigton Heights", "West Harlem", "Hamilton Heights", "Manhattanville", "Morningside Heights", "Central Harlem", "Harlem", "Astor Row", "Sugar Hill", "Marcus Garvey Park", "Le Petit Senegal", "East Harlem", "Upper East Side", "Lenox Hill", "Carnegie Hill", "Yorkville", "Upper West Side", "Lincoln Square", "Morningside Heights", "Washington Heights North", "Washington Heights South", "Inwood", "Inwood Hill Park", "Highbridge Park", "Hamilton Heights", "Central Harlem North", "East Harlem North", "Morningside Heights"))

taxi_nta_mnh_summary_uptown <- group_summarize(taxi_nta_mnh_uptown, zone_dropoff)


taxi_nta_mnh_midtown <- filter(taxi_nta_mnh, zone_dropoff %in% c("Midtown", "Columbus Circle", "Sutton Place", "Rickefeller Center", "Diamond District", "Trutle Bay", "Midtown East", "Midtown", "Tudon City", "Little Brazil" , "Hudson Yards", "Midtown West", "Garment District", "Koreatown", "Murray Hill", "Tenderloin", "Madison Sq", "Clinton West", "East Chelsea", "Flatiron", "Garment District", "Union sq" , "Kips Bay", "Murray Hill", "West Chelsea/Hudson Yards", "Sutton Place/Turtle Bay North", "UN/Turtle Bay South","Sutton Place/Turtle Bay North", "Penn Station/Madison Sq West", "Times Sq/Theatre District", "Gramercy", "Union Sq", "Midtown North", "Midtown South"))

taxi_nta_mnh_summary_midtown <- group_summarize(taxi_nta_mnh_midtown, zone_dropoff)


taxi_nta_mnh_downtown <- filter(taxi_nta_mnh, zone_dropoff %in% c("East Village", "Alphabet City", "TriBeCa/Civic Center", " Little Italy/NoLiTa", "SoHo", "Hudson Sq", " Lower East Side", "Chinatown", "seaport","Two Bridges/Seward Park", "World Trade Center", "Financial District South", "Financial District North", "Battery Park City", "Battery Park", "Greenwich Village North", "Greenwich Village South", "Meatpacking/West Village West", "West Village", "Seaport", "Lower East Side", "Little Italy/NoLiTa"))

taxi_nta_mnh_summary_downtown <- group_summarize(taxi_nta_mnh_downtown, zone_dropoff)


taxi_nta_mnh_uppereast <- filter(taxi_nta_mnh, zone_dropoff %in% c("Yorkville East", "Upper East Side North", "Upper East Side South", "Yorkville West", "East Harlem South", "Lenox Hill West", "Lenox Hill East"))

taxi_nta_mnh_summary_uppereast <- group_summarize(taxi_nta_mnh_uppereast, zone_dropoff)


taxi_nta_mnh_upperwest <- filter(taxi_nta_mnh, zone_dropoff %in% c( "Lincoln Square East", "Lincoln Square West","Bloomingdale",  "Upper West Side North", "Upper West Side South", "Manhattan Valley", "Bloomingdale"))
```

```{r}
taxi_nta_summary <- group_summarize(taxi_nta, zone_dropoff)

taxi_nta_summary$tip_bin <- "0-15"
taxi_nta_summary$tip_bin[between(taxi_nta_summary$tip_percent_mean, .15, .20)] <- "15-20"
taxi_nta_summary$tip_bin[between(taxi_nta_summary$tip_percent_mean, .20, .25)] <- "20-25"
taxi_nta_summary$tip_bin[between(taxi_nta_summary$tip_percent_mean, .25, Inf)] <- "25+"

taxi_nta_mnh_summary$tip_bin <- "0-15"
taxi_nta_mnh_summary$tip_bin[between(taxi_nta_mnh_summary$tip_percent_mean, .15, .20)] <- "15-20"
taxi_nta_mnh_summary$tip_bin[between(taxi_nta_mnh_summary$tip_percent_mean, .20, .25)] <- "20-25"
taxi_nta_mnh_summary$tip_bin[between(taxi_nta_mnh_summary$tip_percent_mean, .25, Inf)] <- "25+"

zone_pickup_summary$tip_bin <- "0-15"
zone_pickup_summary$tip_bin[between(zone_pickup_summary$tip_percent_mean, .15, .21)] <- "15-21"
zone_pickup_summary$tip_bin[between(zone_pickup_summary$tip_percent_mean, .21, .25)] <- "21-25"
zone_pickup_summary$tip_bin[between(zone_pickup_summary$tip_percent_mean, .25, Inf)] <- "25+"

departure_laguardia_summary <- group_summarize(departure_laguardia, zone_pickup)
departure_laguardia_mnh_summary <- group_summarize(departure_laguardia_mnh, zone_pickup)

departure_laguardia_summary$tip_bin <- "0-15"
departure_laguardia_summary$tip_bin[between(departure_laguardia_summary$tip_percent_mean, .15, .20)] <- "15-20"
departure_laguardia_summary$tip_bin[between(departure_laguardia_summary$tip_percent_mean, .20, .21)] <- "20-21"
departure_laguardia_summary$tip_bin[between(departure_laguardia_summary$tip_percent_mean, .21, .25)] <- "21-25"
departure_laguardia_summary$tip_bin[between(departure_laguardia_summary$tip_percent_mean, .25, Inf)] <- "25+"

departure_laguardia_mnh_summary$tip_bin <- "0-15"
departure_laguardia_mnh_summary$tip_bin[between(departure_laguardia_mnh_summary$tip_percent_mean, .15, .20)] <- "15-20"
departure_laguardia_mnh_summary$tip_bin[between(departure_laguardia_mnh_summary$tip_percent_mean, .20, .21)] <- "20-21"
departure_laguardia_mnh_summary$tip_bin[between(departure_laguardia_mnh_summary$tip_percent_mean, .21, .25)] <- "21-25"
departure_laguardia_mnh_summary$tip_bin[between(departure_laguardia_mnh_summary$tip_percent_mean, .25, Inf)] <- "25+"
```

```{r}

```

According to an article by Trip Savvy, drivers in New York City should be tipped an average of 10 to 20 percent of the total cost of a ride. Most of the time, this amount is automatically calculated based on the cost of the fare. However, this value is not fixed and can change if the expectations of the passenger are not meet by the quality of the overall ride. 

To look deeper into what variables impact the total percentage tip for a New York City taxi ride, the taxi dataset was analyzed. The data show that in contrast with the advice, the overall percentage of tip for taxi rides in New York City is above 20 percent. Furthermore, percentage tip has a negative correlation with the total cost of the trip and can vary by borough location.
 

The dataset was analyzed in three distinct ways; in a macro level, by borough level, and by looking at rides from La Guardia Airport. First, at a macro level, the average percentage tip for taxi rides in the entire New York City area was off 24.50 percent of the total trip. There is an overall decreasing trend of the percentage tip as the total cost of the ride increases. A possible explanation for this is customers paying with cash for rides under 20 dollars round up their tip to avoid change. Consequently,  for lower total fare amount, rounding up to the nearest dollar has a more significant influence on the overall percentage tip than for more substantial fares. 

```{r}
model <- lm_basic(tip_percent*100 ~ 1 +borough_dropoff +fare_amount  , data = taxi_nta)
 reg_table(model,level= 0.95) 
```

On the following graph, the previously mentioned trend is visible. When the total fare is under 20 dollars, the percentage tip is primarily  20 to 25 percent. Moreover, when the entire fare is between 20 to 40 dollars, the most common tip is between 15 to 20 percent. (Tips that are higher than 25 percent of the total fare are concentrated around 50 dollars and 70 dollar rides. Again, this trend could be driven by people rounding up their tips to 15 dollars and 20 dollars accordingly.)

```{r}

ggplot(taxi_nta_summary, aes(fare_amount_mean))+
  geom_density(aes(fill= tip_bin), alpha= 0.4)+
  theme_minimal()+
  labs(title= "Distribution of Tips for NYC Rides", x="Fare Amount", fill= "Percentage Tip")



```

Next, the percentage tip by borough was analyzed by looking at the five boroughs in New York City; Bronx, Brooklyn, Manhattan, Queens, and Staten Island. Out of these, Manhattan's overall tip is the lowest with 0.86 percent less than the average percentage tip for the city. On the other hand, Queens and Staten Island both have significantly higher percentage tips of 2.81 percent and 8.25 percent above the city's mean. 

Furthermore, Manhattan was analyzed since it has a smaller mean percentage tip and has significantly more rides than the other zones. The following plot displays all taxi rides that were dropped off in the Manhattan area by zone.  As displayed in yellow, the mean percentage tip is 20 to 25 percent for rides that are lower than 17.5 dollars. 


```{r}
taxi_nta_mnh_summary$tip_bin <- "0-15"
taxi_nta_mnh_summary$tip_bin[between(taxi_nta_mnh_summary$tip_percent_mean, .15, .20)] <- "15-20"
taxi_nta_mnh_summary$tip_bin[between(taxi_nta_mnh_summary$tip_percent_mean, .20, .25)] <- "20-25"
taxi_nta_mnh_summary$tip_bin[between(taxi_nta_mnh_summary$tip_percent_mean, .25, Inf)] <- "25+"

ggplot(taxi_nta_mnh_summary, aes( fare_amount_mean, tip_amount_mean)) +
  geom_point(aes(size=n, color = tip_bin)) +
  scale_color_viridis(discrete = TRUE) +
  theme_minimal()+
  geom_smooth(method="lm", color= "gray")+
  labs(title= "Distribution of Tips for All Manhattan Zone Dropoffs", x="Fare Amount", y= "Tip Amount" , legend= "Percentage Tip")+
  annotate("text", x= 12.46, y=2.53, label= "Chinatown", size= 3)+
   annotate("text", x= 12.44, y=2.3, label= "Lower East Side", size =3)
```

Finally, by looking only at rides leaving La Guardia Airport with a Manhattan destination, the percentage tip by neighborhood and zone was analyzed. This limited the impact of other variables that could indirectly be influencing the percentage of tip. By segmenting the Manhattan area into five neighborhoods, we found that the Upper West Side has the highest percentage tip. However, it has a total median fare of $32.06 which is not the lowest in the area. The actual lowest mean fare for rides coming from La Guardia is for Uptown Manhattan (which makes sense given the proximity of the neighborhood to the airport). However, Uptown has a percentage tip of 22.19 percent, which is the lowest among the Manhattan neighborhoods. 

The graph below shows the distribution of tips for rides from La Guardia Airport into zones in Manhattan. Battery Park, located at the southern tip of the Manhattan Island, stands out since it has the highest average fare price for all zones. On the other hand, East Harlem North has the lowest average fare price and the highest percentage of tip. 


```{r}
ggplot(zone_pickup_summary, aes( fare_amount_mean, tip_amount_mean)) +
  geom_point(aes(size=n, color=tip_percent_mean)) +
  annotate("text", x = 41.8, y = 9.16, label = "Battery Park" , size= 3) +
  annotate("text", x = 21.3, y = 5.7, label = "East Harlem North", size=3 ) + 
  scale_color_viridis() +
  theme_minimal() +
  labs(title= "Distribution of Tips for Rides From La Guardia to Manhattan", x="Fare Amount", y= "Tip Amount" , legend= "Percentage Tip", fill= "Tip Percent")
```


Much more time is needed to understand what drives tips for New York City taxi rides. However, from this short analysis we can conclude that in most cases, the higher the cost of the ride the lower the percentage of tips is. Furthermore, a taxi dropping of customers in Queens, and Staten Island tend to receive higher tips. In Manhattan, the Upper West Side might provide higher tips (specially coming from La Guardia Airport). Yet, La Guardia rides might have a large percentage of tourists that do not represent the New York population. 





















